---
title: "Math 247: Exploration 2.4"
author: "Your name here "
header-includes:
  - \usepackage{float}
  - \floatplacement{figure}{H}
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE,message=FALSE,warning=FALSE, fig.align = 'center', out.width="80%" )
library(tidyverse)
library(ISIwithR)
library(mosaic)
#library(readr)
```

## Backpack Weights

Carrying a heavy backpack can be a source of “chronic, low-level trauma,” and can cause long-lasting shoulder, neck, and back pain. The American Academy of Orthopedic Surgeons recommends that a **backpack not weigh more than 10 to 15% of the wearer’s body weight**. As the problem of overheavy backpacks has received more attention recently among primary and secondary school students, a student group decided to investigate the issue with students at their own university (approximately 20,000 enrolled students). Twenty-five students were sampled from each of four locations across the campus (i.e. a total of 100), during the course of four days, and four different times of the day (early morning, lunch time, evening, late afternoon). Scales were used to measure backpack weight to the nearest pound, and students were asked to report their body weight on a survey. One researcher was the designated talker and introduced the survey to the participants (survey questions included major, year in school, whether have back problems, and gender). Another person in the group was in charge of reading the scale, and another in charge of gathering the surveys.

*Based on the study description, we will investigate the following research question: "Do University students carry backpacks that weigh more than 10% of the owner’s body weight?"*

**STEP 1:** Ask a research question.

> 1.  Identify the population and parameter of interest. What symbol could you use for this parameter? (Hint: for now work with average percent of body weight)

**STEP 2:** Design a study and collect data.

These data have already been collected, so now let us think about the students’ data collection process.

> 2.  Was this a random sample? Is it likely to be representative of the population you identified in #2?

> 3.  What other steps did the students take to minimize non-sampling concerns?

**STEP 3:** Explore the data.

The data in `Backpack.csv` file posted on Canvas. Upload the `Backpack.csv` file to R studio.

Then,read this `Backpack.csv` file into R Markdown with the commands included in the R chunk below.

*Hint*: to get the correct path to include between the quotation marks, locate the csv file in the *Files* tab, click on the down arrow to the right of *More(Blue Wheel Icon)* and choose *Copy folder path to Clipboard* from the drop-down menu.

In my case the path copied to Clipboard is *"\~/Documents/Math 247 Materials/Math 247 Fall 2024/Explorations"*. Paste this path before `/backpack.csv` within the quotation marks in R code chunk below. Thus, *my full path* is :*"\~/Documents/Math 247 Materials/Math 247 Fall 2024/Explorations/backpack.csv".*

Use `head` command to make sure that the file is read correctly.

Remove **eval=FALSE** after running the code.

```{r, reading backpack data, eval=FALSE}
library(readr)
Backpack <- read_csv("~/Documents/Math 247 Materials/Math 247 Fall 2025/Explorations/Backpack.csv")
head(Backpack, n=2)
```

The data in `Backpack.csv` contain several variables of interest, including `Backpack weight`. But the recommendation was stated in terms of “percentage of person’s weight,” so we first convert this variable into the `percentage` of body weight for each individual in the dataset. Remove **eval=FALSE** after running the code.

```{r, creating percentage variable, eval=FALSE}
library(tidyverse)
Backpack <- Backpack %>% 
  mutate(percentage=(BackpackWt/BodyWt)*100) %>% 
  mutate(percentage=round(percentage, digits = 1)) %>% 
  relocate(percentage)
head(Backpack,2)
```

Let's create the distribution of the percentages for this sample of 100 students.Remove **eval=FALSE** after running the code.

```{r, visualizing the distribution of percentages for this sample, eval=FALSE}
histogram(~percentage, data = Backpack)
```

> 4.  Describe the distribution of the percentages for this sample of 100 students. Remember to focus on shape, center, and variability, and any unusual observations or outliers.

Remove **eval=FALSE** after running the code.

```{r, calculating the sample mean and sample standard deviation, eval=FALSE}
favstats(~percentage, data = Backpack)[c("mean","sd")]
```

The `favstats` function above reports the mean and the standard deviation. Of course, another way to measure the “center” of the distribution is the median.

> 5.  Predict how the median percentage will compare to the mean percentage, keeping in mind that the median will divide the dataset in half, 50 students below the median and 50 students above the median.

> 6.  Using the R code below report the median. Was your conjecture correct? Explain to someone why the mean and median have this relationship for this dataset.

Remove **eval=FALSE** after running the code.

```{r, calculating sample median, eval=FALSE}
favstats(~percentage, data = Backpack)["median"]
```

> 7.  The largest percentage in this dataset is 18.1%. Suppose this had been incorrectly recorded as 181%! How will the mean change? How will the median change?

*For this dataset, the difference between the mean and median is not large, but we still might consider the median a more appropriate statistic to focus on, as it would represent a **typical** percentage, rather than the mean which is being pulled to the right by a few large percentages. The problem is, all the theory you saw in Sections 2.2 and 2.3 only applies to means. So how can we get a sense for how much sample-to-sample variation there is in the sample medians?*

**STEP 4.** Draw inferences beyond the data.

*Is it plausible that this sample came from a population with a population median of 10%?*

> 8.  State appropriate null and alternative hypotheses, in words, about the population median for this research question. (Although inference for the median will follow the same logic as inference for the mean, we will not bother to define symbols for the sample or population median, we will use words instead.)

Null hypothesis: $H_0: population  \  median = 10\%$

Alternative hypothesis:

> 9.  Outline a simulation strategy for this null hypothesis:

Population:

Parameter:

Sample:

Statistic:

*In Section 2.3, we created hypothetical populations from which to sample. We wanted the population to have similar characteristics as the observed sample, but also assuming the null hypothesis to be true. We could generate a population of 20,000 percentages, with a population median of 10, but what do we assume about the shape and the standard deviation of this population? Do they matter? Does the population size?*

*One method for generating a hypothetical population is to make many copies of the observed sample. In the Exercises in Section 2.2, we suggested that as long as the population is more than 20 times the size of the sample, then the actual population size does not matter.This process, sampling from a large number of copies of the sample distribution, is equivalent to sampling from the original sample with replacement, a technique known as bootstrapping.*

Let's use R to generate 1000 bootstrap samples with replacement of size 100 from the original data and calculate medians of each of 1000 samples. First, install a package `infer` using `Packages->Install` in the bottom right pane.

Remove **eval=FALSE** after running the code.

```{r, generating 1000 bootstrap samples, eval=FALSE}
library(infer)
set.seed(356)
# Generate bootstrap distribution of medians
bootobject <- Backpack %>%
  # Specify the variable of interest
  specify(response = percentage) %>%  
  # Generate 1000 bootstrap samples
  generate(reps = 1000, type = "bootstrap") %>% 
  # Calculate the median of each bootstrap sample
 calculate(stat = "median")
head(bootobject, n=3)
```

*The distributions you just generated are often called “bootstrap distributions.”*

> 10. Let's use R to create a distribution of medians based on sampling with replacement from the original sample. How would you characterize the shape, center, and variability of this distribution?

Remove **eval=FALSE** after running the code.

```{r, visualiaing distribution of medians based on sampling with replacement, eval=FALSE}
histogram(~stat,
          xlab="sample median",
          main="Bootstrap Distribution of sample median",
          cex.main=0.6,data = bootobject)
```

```{r, calculating the mean and SD of the distribution of sample medians, eval=FALSE}
favstats(~stat, data = bootobject)[c("mean","sd")]
```

Notice that the distribution of medians does not follow a nice smooth bell-shaped curve like the distribution of means, but now we are able to see how the distribution behaves. This process, sampling from a large number of copies of the sample distribution, is equivalent to sampling from the original sample with replacement, a technique known as *bootstrapping*. The distributions you just generated are often called “bootstrap distributions.”

**Definition**

*Bootstrapping* randomly samples n observations from the original sample of n observations with replacement to estimate the sample-to-sample variation in a statistic. This is equivalent to considering the “population” from which you are sampling to be a very large number of copies of the sample.

**Key Idea**: Bootstrapping allows you to estimate the sample to sample variation in any statistic.

## Null Distribution?

*We cannot use the bootstrap distribution directly to estimate a p-value* because it will be centered at the observed sample statistic rather than at the hypothesized value for the parameter. However, the bootstrap distribution does tell us about the sample-to-sample variation in the statistic. In fact, with quantitative data, we can simply shift the “population distribution” over to center at the hypothesized parameter value without changing the shape or variability of the distribution.

> 11. Let's use R to recenter the bootstrap distribution so that the center is at the value of the null hypothesis, i.e. 10. Remove **eval=FALSE** after running the code.

```{r, recentering the bootstrap distribution at the value of the null, eval=FALSE}
bootobject_c<-bootobject %>% 
  mutate(stat_centered=stat+2.9)
```

Alternatively, we could regenerate the bootstrap samples using the same seed as for the generation of the `bootobject` but tell R to center the distribution at the value of 10 directly.

```{r, another way to re-center the bootstrap distribution at the value of the null, eval=FALSE}
set.seed(356)
# Generate 15000 bootstrap samples centered at null
bootobject_centered <- Backpack %>%
  specify(response = percentage) %>%
  # center the distribution at the value of 10
  hypothesize(null = "point", med =10) %>%
  # Generate 1000 bootstrap samples
  generate(reps = 1000, type = "bootstrap") %>% 
  # Calculate the median of each bootstrap sample
  calculate(stat = "median")
head(bootobject_centered, n=2)
```

> 12. Let's visualize the distribution of centered bootstrap samples. How would you characterize the shape, center, and variability of this population compared to the non-centered bootstrap distribution?

Remove **eval=FALSE** after running the code.

```{r, visualize the distribution of centered bootstrap samples, eval=FALSE}
histogram(~stat_centered, 
           xlab="sample median",
          main="Centered Bootstrap Distribution of sample median",
          cex.main=0.65,
          data = bootobject_c)
```

Alternatively, use the code that is based on `bootobject_centered`

```{r, visualize the distribution of centered bootstrap samples - another way, eval=FALSE}
# 7.1 is the sample median in the original data
histogram(~stat,
          xlab="sample median",
          main="Centered Bootstrap Distribution of sample median",
          cex.main=0.6,
          data = bootobject_centered)
```

```{r, center and spread of the distribution of centered bootstrap samples, eval=FALSE}
favstats(~stat, data = bootobject_centered)[c("mean","sd")]
```

> 13. Is it plausible that the observed sample of 100 percentages came from a population where the median percentage was 10? Explain your reasoning.

**STEP 5**: Formula conclusions.

> 14. Write a paragraph summarizing your conclusions from this study. Be sure to discuss both what you learned from the sample and what you believe to be plausible about the population.

**STEP 6**: Look back and ahead.

> 15. Return to your critique of the study design. If you were to carry out such a study on your own campus, discuss what you would do differently and why.

## Summary

In many situations the mean will not be the primary statistic of interest. When exploring other statistics, like a median, we do not have a theory-based result for how the sampling distribution will behave. One method for exploring the sampling distribution of the statistic is bootstrapping. The idea behind bootstrapping is to sample from a population that is constructed from infinitely many copies of the observed sample. You explored this by randomly sampling, with replacement, from the observed sample. The key information from this bootstrap distribution is the standard deviation. This tells us how much sample-to-sample variation we can expect in the statistic from the random sampling process alone. Knowing the standard deviation gives you a way of measuring whether the observed statistic is "far" from the hypothesized value.

*Make sure to Remove all **eval=FALSE** before knitting the document.*
